### 2024-04-30 misc/expo/entropy.tex
Ahoy

\edef\lemfivetwo{\the\sectcount.\the\thmcount}
\parenproclaim Lemma {\advthm} ({\rm\ref{pfr2023},} Lemma 5.2).
Let $X$, $Y$, and $Z$ be random variables taking values in an abelian group,
with $Y$ and $Z$ independent. Then
$$\eqalign{
\dd\{X, Y-Z\} - \dd\{X, Y\} &\le {\Eta\{Y-Z\} - \Eta\{Y\}\over 2} \cr
&= {1\over 2} \dd\{Y, Z\} + {1\over 4}\Eta\{Z\} - {1\over 4}\Eta\{Y\}\cr
}$$
and
$$\eqalign{
\dd\{X; Y\given Y-Z\} - \dd\{X, Y\} &\le {\Eta\{Y-Z\} - \Eta\{Z\}\over 2} \cr
&= {1\over 2} \dd\{Y, Z\} + {1\over 4}\Eta\{Z\} - {1\over 4}\Eta\{Z\}.\cr
}$$

@@

\proof From the fact that any two of the $T_j$ determine the full triple $(T_1, T_2, T_3)$,
we have
$$\Eta\{T_1, T_2\} = \Eta\{T_1, T_3\} = \Eta\{T_2,T_3\} = \Eta\{T_1, T_2, T_3\}.$$
Then since $T_1 + T_2 = T_3$ we may apply Lemma~{\lematwopfr} with
$(X,Y,Z) = (T_1, T_2, T_3)$ to bound
$$\eqalign{
\sum_{t_3\in \FF_2^n} \pr\{T_3 &= t_3\} \dd\bigl\{ (T_1\given T_3 = t_3);
(T_2\given T_3 = t_3)\bigr\} \cr
&\le 2\II\{T_1:T_2\} + 2\Eta\{T_3\} - \Eta\{T_1,T_2\} \cr
&= 2\Eta\{T_1\} + 2\Eta\{T_2\} + 2\Eta\{T_3\} - 3\Eta\{T_1,T_2\} \cr
&= \II\{T_1,T_2\} + \II\{T_1, T_3\} + \II\{T_2,T_3\} \cr
&= \delta.
}$$
Letting $Z$ be a random variable independent of both $X_1^*$ and $X_2^*$, we apply Lemma~{\lemfiveone}
with $(X,Z,Y,W) = (X_1^*, Z, T_1, T_3)$ to obtain
$$\eqalign{
\sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\bigl(\dd\bigl\{X_1^*;(T_1\given T_3&=t_3)\bigr\}-\dd\{X_1^*, X_1\}\bigr) \cr
&= \dd\{ X_1^*, T_1\given T_3\} - \dd\{X_1^*, X_1\} \cr
&\le \dd\{X_1^*, T_1\} + {1\over 2}\II\{T_1:T_3\} - \dd\{X_1^*, X_1\}, \cr
}$$
and applying the same lemma with $(X,Z,Y,W) = (X_2^*, Z, T_2, T_3)$ yields
$$\eqalign{
\sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\bigl(\dd\bigl\{X_2^*;(T_2\given T_3&=t_3)\bigr\}-\dd\{X_2^*, X_2\}\bigr) \cr
&\le \dd\{X_2^*, T_2\} + {1\over 2}\II\{T_2:T_3\} - \dd\{X_2^*, X_2\}. \cr
}$$
Putting the three observations together, we have
$$\eqalign{
&\sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\psi\bigl\{(T_1\given T_3=t_3);(T_2\given T_3=t_3)\bigr\} \cr
&\hskip10pt = \sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\Bigl(\dd\bigl\{(T_1\given T_3=t_3);(T_2\given T_3=t_3)\bigr\} \cr
&\hskip103pt + \eta\dd\bigl\{X_1^*;(T_1\given T_3=t_3)\bigr\}+ \eta\dd\bigl\{X_2^*;(T_2\given T_3=t_3)\bigr\} \cr
&\hskip192pt - \eta\dd\{X_1^*, X_1\} - \eta\dd\{X_2^*, X_2\} \Bigr) \cr
&\hskip10pt \le \delta + \eta\bigl( \dd\{X_1^*, T_1\} - \dd\{X_1^*, X_1\}+\dd\{X_2^*, T_2\}-\dd\{X_2^*, X_2\}\cr
&\hskip180pt 
+ {1\over 2} \II\{T_1 : T_3\} + {1\over 2}\II\{T_2:T_3\} \bigr), \cr
}$$
Choosing some $t_3$ in the support of $T_3$ such that the value of $\phi$ inside
the sum is minimal and then setting $T_{1,3}' = (T_1\given T_3 = t_3)$ and
$T_{2,3}' = (T_2\given T_3 = t_3)$, we have
$$\eqalign{
\psi\bigl\{T_{1,3}', T_{2,3}'\bigr\} \le \delta + \eta\bigl( \dd\{X_1^*, T_1\} - \dd\{X_1^*, X_1\}
&+ \dd\{X_2^*, T_2\} - \dd\{X_2^*, X_2\} \cr
& + {1\over 2} \II\{T_1 : T_3\} + {1\over 2}\II\{T_2:T_3\} \bigr)\cr
}$$
We can now repeat this for all six permutations $(\alpha, \beta, \gamma)$ of $(1,2,3)$ and average the
corresponding bounds for $(T_\alpha, T_\beta, T_\gamma)$ to obtain
$$\eqalign{
{1\over 6} \sum_{(\alpha,\beta,\gamma)\in {\frak S}_3}
\psi\bigl\{T_{\alpha,\gamma}', T_{\beta,\gamma}'\bigr\}
\le \delta &- \eta\dd\{X_1^*, X_1\} - \eta\dd\{X_2^*, X_2\} \cr
&+ {\eta\over 6}\sum_{(\alpha,\beta,\gamma)\in {\frak S}_3}
\bigl( \dd\{X_1^*, T_\alpha\} + \dd\{X_2^*, T_\beta\} \cr
&\hskip68pt+ {1\over 2} \II\{T_\alpha : T_\gamma\} + {1\over 2}\II\{T_\beta:T_\gamma\} \bigr).\cr
}$$
Since each of $\{1,2,3\}$ appears twice as $\alpha$ and twice as $\beta$, we have
$$\sum_{(\alpha,\beta,\gamma)\in {\frak S}_3} \bigl(\dd\{X_1^*, T_\alpha\} + \dd\{X_2^*, T_\beta\}\bigr)
= 2\sum_{i=1}^2\sum_{j=1}^3 \dd\{X_i^*, T_j\}$$
and
$$\sum_{(\alpha,\beta,\gamma)\in {\frak S}_3}
\biggl({1\over 2} \II\{T_\alpha : T_\gamma\} + {1\over 2}\II\{T_\beta:T_\gamma\}\biggr) = 2\delta.$$
Consequently, the average of
$\psi\bigl\{T_{\alpha,\gamma}', T_{\beta,\gamma}'\bigr\}$
over all permutations $(\alpha,\beta,\gamma)$ is bounded above by
$$\biggl(1+{\eta\over 3}\biggr)\delta
- \eta\dd\{X_1^*, X_1\} -\eta\dd\{X_2^*, X_2\} + {\eta\over 3}\sum_{i=1}^2\sum_{j=1}^3 \dd\{X_i^*, T_j\},$$
so the result follows by letting $(T_1', T_2')$ equal the pair $(T_{\alpha,\gamma}', T_{\beta,\gamma}')$
for the choice of $(\alpha,\beta,\gamma)$ that minimises
$\psi\bigl\{T_{\alpha,\gamma}', T_{\beta,\gamma}'\bigr\}$.\slug

@@

\proof Let $k=\dd\{X_1, X_2\}$. Let $X_1$ and $\bar X_1$ be copies of $X_1$ and $X_2$ and $\bar X_2$ copies
of $X_2$ such that $X_1$, $\bar X_1$, $X_2$, and $\bar X_2$ are all independent. Let $I_1$ and $I_2$
be as in the previous section, and let
$$I_3 = \II\{\bar X_1 + X_2 : X_1 + \bar X_1\given X_1 + X_2 + \bar X_1 + \bar X_2\},$$
so that $I_3 = I_2$ (which follows from interchanging the r\^oles of $X_1$ and $\bar X_1$).
For brevity of notation, let
$$U = X_1+X_2,\qquad V=\bar X_1 + X_2,\qquad W = X_1 + \bar X_1,$$
and
$$S = X_1 + X_2 + \bar X_1 + \bar X_2.$$
Then
$$I_1 = \II\{ U : V\given S\},\qquad I_2 = \II\{W:U\given S\},\qquad\hbox{and}\qquad I_3 = \II\{V:W\given S\}.$$
Lemma~{\lemionebound} gave us
$$I_3 = I_2 \le 2\eta k + {2\eta(2\eta k-I_1)\over 1-\eta},$$
so setting
$$\bar\delta = \II\{U:V\given S\} + \II\{W:U\given S\} + \II\{V:W\given S\},$$
we have
\edef\eqbardeltabound{\the\eqcount}
$$\eqalign{
\bar \delta &\le I_1+4\eta k+{4\eta(2\eta k - I_1)\over 1-\eta} \cr
&= {I_1 - \eta I_n + 4\eta k - 4\eta^2 k + 6\eta^2 k - 4\eta I_1\over 1-\eta} \cr
&= 6\eta k + {(1-5\eta)I_1 + 4\eta k + 4\eta^2 k - 6\eta k + 6\eta^2 k \over 1-\eta} \cr
&= 6\eta k + {(1-5\eta)I_1 - 2\eta k + 10\eta^2 k \over 1-\eta} \cr
&= 6\eta k - {1-5\eta \over 1-\eta}(2\eta k - I_1) \cr
}\adveq$$

Now we peform invocations of Lemma~{\lemsevenone} to obtain bounds on various Ruzsa distances. Setting
$(X,Y,Z,W) = (X_1^*, X_1, X_2, \bar X_1+\bar X_2)$, gives
$$\eqalign{
\dd\{X_1^*; U\given S\} - \dd\{X_1^*, X_1\}
&\le {1\over 2}\bigl( \Eta\{S\} + \Eta\{U\} - \Eta\{X_1\} - \Eta\{\bar X_1+\bar X_2\}\bigr)\cr
&= {1\over 2} \bigl( \Eta\{S\} - \Eta\{X_1\}\bigr),
}$$
where in the second line we used the fact that
$$\Eta\{U\} = \Eta\{X_1 + X_2\} = \Eta\{\bar X_1 + \bar X_2\}.$$
Similarly, we have
$$\eqalign{
\dd\{X_2^*; U\given S\} - \dd\{X_2^*, X_2\} &= {1\over 2} \bigl( \Eta\{S\} - \Eta\{X_2\}\bigr), \cr
\dd\{X_1^*; V\given S\} - \dd\{X_1^*, X_1\} &= {1\over 2} \bigl( \Eta\{S\} - \Eta\{X_1\}\bigr), \cr
\dd\{X_2^*; V\given S\} - \dd\{X_2^*, X_2\} &= {1\over 2} \bigl( \Eta\{S\} - \Eta\{X_2\}\bigr), \cr
}$$
and
$$\dd\{X_1^*, W\given S\} - \dd\{X_1^*, X_1\}
= {1\over 2} \bigl( \Eta\{S\} + \Eta\{W\} - \Eta\{X_1\} - \Eta\{W'\}\bigr),$$
where $W' = X_2 + \bar X_2$. To address the asymmetry in the last bound, we note that for any fixed value
$s$ taken by $S$, we have $W' = W + s$, so $\dd\{X_2^*; W\given S\} = \dd\{X_2^*; W'\given S\}$. Now
applying Lemma~{\lemsevenone} to $W'$ yields
$$\dd\{X_1^*; W\given S\} - \dd\{X_1^*, X_1\}
= {1\over 2} \bigl( \Eta\{S\} + \Eta\{W\} - \Eta\{X_1\} - \Eta\{W'\}\bigr).$$
The sum of these six inequalities is
\edef\eqsixsum{\the\eqcount}
$$\sum_{i=1}^2 \sum_{A\in \{U,V,W\}} \bigl( \dd\{X_i^*; A\given S\} - \dd\{X_i^*, X_i\}\bigr)
\le 3\Eta\{S\} - {3\over 2} \Eta\{X_1\} - {3\over 2} \Eta\{X_2\}.\adveq$$
The second part of Lemma~{\lemionebound} states that
$$\Eta\{S\} \le {1\over 2} \Eta\{X_1\} + {1\over 2} \Eta\{X_2\} + (2+\eta)k - I_1,$$
which can be plugged into~\refeq{\eqsixsum} to give
\edef\eqdistancebounds{\the\eqcount}
$$\eqalign{
\sum_{i=1}^2 \sum_{A\in \{U,V,W\}} \bigl( \dd\{X_i^*; A\given S\} - \dd\{X_i^*, X_i\}\bigr)
&\le 3(2+\eta)k - 3I_1 \cr
&\le 6k -3\eta k + 6\eta k - 3I_1 \cr
&\le (6-3\eta)k + 3(2\eta k - I_1).\cr
}\adveq$$

Let $\psi$ be defined as in Lemma~{\lemseventwo}. By Lemma~{\lemionehelper}, $k\le \psi\{T_1', T_2'\}$
for all random variables $T_1'$ and $T_2'$, so
for any random variables $(T_1, T_2, T_3)$, the pair $(T_1',T_2')$ furnished by Lemma~{\lemseventwo}
satisfies
$$k\le \psi\{T_1', T_2'\} \le \biggl(1+{\eta\over 3}\biggr)\delta
+ {\eta\over 3}\sum_{i=1}^2\sum_{j=1}^3 \bigl(\dd\{X_i^*, T_j\}-\dd\{X_i^*, X_i\}\bigr).$$
In particular, since $(U+V+W\given S=s)$ is identically zero for all possible values $s$ of $S$,
we can set
$$T_1 = (U\given S=s),\qquad T_2 = (V\given S=s), \qquad\hbox{and}\qquad T_3 = (W\given S=s)$$
and then average the above inequality over all $s\in \FF_2^n$, weighted by $\pr\{S=s\}$ to obtain
$$k\le \biggl(1+{\eta\over 3}\biggr)\bar\delta
+ {\eta\over 3}\sum_{i=1}^2\sum_{A\subseteq\{U,V,W\}} \bigl(\dd\{X_i^*, A\given S\}-\dd\{X_i^*, X_i\}\bigr).$$
We now substitute the upper bounds~\refeq{\eqbardeltabound} and~\refeq{\eqdistancebounds} to get
$$\eqalign{
k &\le \biggl(1+{\eta\over 3}\biggr)\biggl(6\eta k - {1-5\eta \over 1-\eta}(2\eta k - I_1)\biggr)
+ {\eta\over 3}\biggl( (6-3\eta)k + 3(2\eta k - I_1)\biggr) \cr
&= (8\eta - \eta^2)k + \biggl(\eta-\Bigl(1+{\eta\over 3}\Bigr){1-5\eta\over 1-\eta} \biggr)
(2\eta k - I_1).\cr
}$$
With the choice $\eta = 1/9$,
$$\eta-\Bigl(1+{\eta\over 3}\Bigr){1-5\eta\over 1-\eta} = -{11\over 27} \le 0$$
and $2\eta k - I_1\ge 0$ by Lemma~{\lemionebound}, hence one concludes that
$$ k \le (8\eta + \eta^2)k = {73\over 81}k,$$
which is nonsense unless $k=0$.\slug

>>>>

\edef\lemfivetwo{\the\sectcount.\the\thmcount}
\parenproclaim Lemma {\advthm} ({\rm\ref{pfr2023},} Lemma 5.2).
Let $X$, $Y$, and $Z$ be random variables taking values in an abelian group,
with $Y$ and $Z$ independent. Then
$$\eqalign{
\dd\{X, Y-Z\} - \dd\{X, Y\} &\le {\Eta\{Y-Z\} - \Eta\{Y\}\over 2} \cr
&= {1\over 2} \dd\{Y, Z\} + {1\over 4}\Eta\{Z\} - {1\over 4}\Eta\{Y\}\cr
}$$
and
$$\eqalign{
\dd\{X; Y\given Y-Z\} - \dd\{X, Y\} &\le {\Eta\{Y-Z\} - \Eta\{Z\}\over 2} \cr
&= {1\over 2} \dd\{Y, Z\} + {1\over 4}\Eta\{Y\} - {1\over 4}\Eta\{Z\}.\cr
}$$

@@

\proof From the fact that any two of the $T_j$ determine the full triple $(T_1, T_2, T_3)$,
we have
$$\Eta\{T_1, T_2\} = \Eta\{T_1, T_3\} = \Eta\{T_2,T_3\} = \Eta\{T_1, T_2, T_3\}.$$
Then since $T_1 + T_2 = T_3$ we may apply Lemma~{\lematwopfr} with
$(X,Y,Z) = (T_1, T_2, T_3)$ to bound
$$\eqalign{
\sum_{t_3\in \FF_2^n} \pr\{T_3 &= t_3\} \dd\bigl\{ (T_1\given T_3 = t_3);
(T_2\given T_3 = t_3)\bigr\} \cr
&\le 2\II\{T_1:T_2\} + 2\Eta\{T_3\} - \Eta\{T_1,T_2\} \cr
&= 2\Eta\{T_1\} + 2\Eta\{T_2\} + 2\Eta\{T_3\} - 3\Eta\{T_1,T_2\} \cr
&= \II\{T_1:T_2\} + \II\{T_1:T_3\} + \II\{T_2:T_3\} \cr
&= \delta.
}$$
Letting $Z$ be a random variable independent of both $X_1^*$ and $X_2^*$, we apply Lemma~{\lemfiveone}
with $(X,Z,Y,W) = (X_1^*, Z, T_1, T_3)$ to obtain
$$\eqalign{
\sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\bigl(\dd\bigl\{X_1^*;(T_1\given T_3&=t_3)\bigr\}-\dd\{X_1^*, X_1\}\bigr) \cr
&= \dd\{ X_1^*, T_1\given T_3\} - \dd\{X_1^*, X_1\} \cr
&\le \dd\{X_1^*, T_1\} + {1\over 2}\II\{T_1:T_3\} - \dd\{X_1^*, X_1\}, \cr
}$$
and applying the same lemma with $(X,Z,Y,W) = (X_2^*, Z, T_2, T_3)$ yields
$$\eqalign{
\sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\bigl(\dd\bigl\{X_2^*;(T_2\given T_3&=t_3)\bigr\}-\dd\{X_2^*, X_2\}\bigr) \cr
&\le \dd\{X_2^*, T_2\} + {1\over 2}\II\{T_2:T_3\} - \dd\{X_2^*, X_2\}. \cr
}$$
Putting the three observations together, we have
$$\eqalign{
&\sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\psi\bigl\{(T_1\given T_3=t_3);(T_2\given T_3=t_3)\bigr\} \cr
&\hskip10pt = \sum_{t_3\in\FF_2^n}\pr\{T_3 = t_3\}\Bigl(\dd\bigl\{(T_1\given T_3=t_3);(T_2\given T_3=t_3)\bigr\} \cr
&\hskip103pt + \eta\dd\bigl\{X_1^*;(T_1\given T_3=t_3)\bigr\}+ \eta\dd\bigl\{X_2^*;(T_2\given T_3=t_3)\bigr\} \cr
&\hskip192pt - \eta\dd\{X_1^*, X_1\} - \eta\dd\{X_2^*, X_2\} \Bigr) \cr
&\hskip10pt \le \delta + \eta\bigl( \dd\{X_1^*, T_1\} - \dd\{X_1^*, X_1\}+\dd\{X_2^*, T_2\}-\dd\{X_2^*, X_2\}\cr
&\hskip180pt 
+ {1\over 2} \II\{T_1 : T_3\} + {1\over 2}\II\{T_2:T_3\} \bigr), \cr
}$$
Choosing some $t_3$ in the support of $T_3$ such that the value of $\psi$ inside
the sum is minimal and then setting $T_{1,3}' = (T_1\given T_3 = t_3)$ and
$T_{2,3}' = (T_2\given T_3 = t_3)$, we have
$$\eqalign{
\psi\bigl\{T_{1,3}', T_{2,3}'\bigr\} \le \delta + \eta\bigl( \dd\{X_1^*, T_1\} - \dd\{X_1^*, X_1\}
&+ \dd\{X_2^*, T_2\} - \dd\{X_2^*, X_2\} \cr
& + {1\over 2} \II\{T_1 : T_3\} + {1\over 2}\II\{T_2:T_3\} \bigr)\cr
}$$
We can now repeat this for all six permutations $(\alpha, \beta, \gamma)$ of $(1,2,3)$ and average the
corresponding bounds for $(T_\alpha, T_\beta, T_\gamma)$ to obtain
$$\eqalign{
{1\over 6} \sum_{(\alpha,\beta,\gamma)\in {\frak S}_3}
\psi\bigl\{T_{\alpha,\gamma}', T_{\beta,\gamma}'\bigr\}
\le \delta &- \eta\dd\{X_1^*, X_1\} - \eta\dd\{X_2^*, X_2\} \cr
&+ {\eta\over 6}\sum_{(\alpha,\beta,\gamma)\in {\frak S}_3}
\bigl( \dd\{X_1^*, T_\alpha\} + \dd\{X_2^*, T_\beta\} \cr
&\hskip68pt+ {1\over 2} \II\{T_\alpha : T_\gamma\} + {1\over 2}\II\{T_\beta:T_\gamma\} \bigr).\cr
}$$
Since each of $\{1,2,3\}$ appears twice as $\alpha$ and twice as $\beta$, we have
$$\sum_{(\alpha,\beta,\gamma)\in {\frak S}_3} \bigl(\dd\{X_1^*, T_\alpha\} + \dd\{X_2^*, T_\beta\}\bigr)
= 2\sum_{i=1}^2\sum_{j=1}^3 \dd\{X_i^*, T_j\}$$
and
$$\sum_{(\alpha,\beta,\gamma)\in {\frak S}_3}
\biggl({1\over 2} \II\{T_\alpha : T_\gamma\} + {1\over 2}\II\{T_\beta:T_\gamma\}\biggr) = 2\delta.$$
Consequently, the average of
$\psi\bigl\{T_{\alpha,\gamma}', T_{\beta,\gamma}'\bigr\}$
over all permutations $(\alpha,\beta,\gamma)$ is bounded above by
$$\biggl(1+{\eta\over 3}\biggr)\delta
- \eta\dd\{X_1^*, X_1\} -\eta\dd\{X_2^*, X_2\} + {\eta\over 3}\sum_{i=1}^2\sum_{j=1}^3 \dd\{X_i^*, T_j\},$$
so the result follows by letting $(T_1', T_2')$ equal the pair $(T_{\alpha,\gamma}', T_{\beta,\gamma}')$
for the choice of $(\alpha,\beta,\gamma)$ that minimises
$\psi\bigl\{T_{\alpha,\gamma}', T_{\beta,\gamma}'\bigr\}$.\slug

@@

\proof Let $k=\dd\{X_1, X_2\}$. Let $X_1$ and $\bar X_1$ be copies of $X_1$ and $X_2$ and $\bar X_2$ copies
of $X_2$ such that $X_1$, $\bar X_1$, $X_2$, and $\bar X_2$ are all independent. Let $I_1$ and $I_2$
be as in the previous section, and let
$$I_3 = \II\{\bar X_1 + X_2 : X_1 + \bar X_1\given X_1 + X_2 + \bar X_1 + \bar X_2\},$$
so that $I_3 = I_2$ (which follows from interchanging the r\^oles of $X_1$ and $\bar X_1$).
For brevity of notation, let
$$U = X_1+X_2,\qquad V=\bar X_1 + X_2,\qquad W = X_1 + \bar X_1,$$
and
$$S = X_1 + X_2 + \bar X_1 + \bar X_2.$$
Then
$$I_1 = \II\{ U : V\given S\},\qquad I_2 = \II\{W:U\given S\},\qquad\hbox{and}\qquad I_3 = \II\{V:W\given S\}.$$
Lemma~{\lemionebound} gave us
$$I_3 = I_2 \le 2\eta k + {2\eta(2\eta k-I_1)\over 1-\eta},$$
so setting
$$\bar\delta = \II\{U:V\given S\} + \II\{W:U\given S\} + \II\{V:W\given S\},$$
we have
\edef\eqbardeltabound{\the\eqcount}
$$\eqalign{
\bar \delta &\le I_1+4\eta k+{4\eta(2\eta k - I_1)\over 1-\eta} \cr
&= {I_1 - \eta I_n + 4\eta k - 4\eta^2 k + 8\eta^2 k - 4\eta I_1\over 1-\eta} \cr
&= 6\eta k + {(1-5\eta)I_1 + 4\eta k + 4\eta^2 k - 6\eta k + 6\eta^2 k \over 1-\eta} \cr
&= 6\eta k + {(1-5\eta)I_1 - 2\eta k + 10\eta^2 k \over 1-\eta} \cr
&= 6\eta k - {1-5\eta \over 1-\eta}(2\eta k - I_1) \cr
}\adveq$$

Now we perform invocations of Lemma~{\lemsevenone} to obtain bounds on various Ruzsa distances. Setting
$(X,Y,Z,W) = (X_1^*, X_1, X_2, \bar X_1+\bar X_2)$, gives
$$\eqalign{
\dd\{X_1^*; U\given S\} - \dd\{X_1^*, X_1\}
&\le {1\over 2}\bigl( \Eta\{S\} + \Eta\{U\} - \Eta\{X_1\} - \Eta\{\bar X_1+\bar X_2\}\bigr)\cr
&= {1\over 2} \bigl( \Eta\{S\} - \Eta\{X_1\}\bigr),
}$$
where in the second line we used the fact that
$$\Eta\{U\} = \Eta\{X_1 + X_2\} = \Eta\{\bar X_1 + \bar X_2\}.$$
Similarly, we have
$$\eqalign{
\dd\{X_2^*; U\given S\} - \dd\{X_2^*, X_2\} &\le {1\over 2} \bigl( \Eta\{S\} - \Eta\{X_2\}\bigr), \cr
\dd\{X_1^*; V\given S\} - \dd\{X_1^*, X_1\} &\le {1\over 2} \bigl( \Eta\{S\} - \Eta\{X_1\}\bigr), \cr
\dd\{X_2^*; V\given S\} - \dd\{X_2^*, X_2\} &\le {1\over 2} \bigl( \Eta\{S\} - \Eta\{X_2\}\bigr), \cr
}$$
and
$$\dd\{X_1^*, W\given S\} - \dd\{X_1^*, X_1\}
\le {1\over 2} \bigl( \Eta\{S\} + \Eta\{W\} - \Eta\{X_1\} - \Eta\{W'\}\bigr),$$
where $W' = X_2 + \bar X_2$. To address the asymmetry in the last bound, we note that for any fixed value
$s$ taken by $S$, we have $W' = W + s$, so $\dd\{X_2^*; W\given S\} = \dd\{X_2^*; W'\given S\}$. Now
applying Lemma~{\lemsevenone} to $W'$ yields
$$\dd\{X_2^*; W\given S\} - \dd\{X_2^*, X_2\}
\le {1\over 2} \bigl( \Eta\{S\} + \Eta\{W\} - \Eta\{X_2\} - \Eta\{W'\}\bigr).$$
The sum of these six inequalities is
\edef\eqsixsum{\the\eqcount}
$$\sum_{i=1}^2 \sum_{A\in \{U,V,W\}} \bigl( \dd\{X_i^*; A\given S\} - \dd\{X_i^*, X_i\}\bigr)
\le 3\Eta\{S\} - {3\over 2} \Eta\{X_1\} - {3\over 2} \Eta\{X_2\}.\adveq$$
The second part of Lemma~{\lemionebound} states that
$$\Eta\{S\} \le {1\over 2} \Eta\{X_1\} + {1\over 2} \Eta\{X_2\} + (2+\eta)k - I_1,$$
which can be plugged into~\refeq{\eqsixsum} to give
\edef\eqdistancebounds{\the\eqcount}
$$\eqalign{
\sum_{i=1}^2 \sum_{A\in \{U,V,W\}} \bigl( \dd\{X_i^*; A\given S\} - \dd\{X_i^*, X_i\}\bigr)
&\le 3(2+\eta)k - 3I_1 \cr
&\le 6k -3\eta k + 6\eta k - 3I_1 \cr
&= (6-3\eta)k + 3(2\eta k - I_1).\cr
}\adveq$$

Let $\psi$ be defined as in Lemma~{\lemseventwo}. By Lemma~{\lemionehelper}, $k\le \psi\{T_1', T_2'\}$
for all random variables $T_1'$ and $T_2'$, so
for any random variables $(T_1, T_2, T_3)$, the pair $(T_1',T_2')$ furnished by Lemma~{\lemseventwo}
satisfies
$$k\le \psi\{T_1', T_2'\} \le \biggl(1+{\eta\over 3}\biggr)\delta
+ {\eta\over 3}\sum_{i=1}^2\sum_{j=1}^3 \bigl(\dd\{X_i^*, T_j\}-\dd\{X_i^*, X_i\}\bigr).$$
In particular, since $(U+V+W\given S=s)$ is identically zero for all possible values $s$ of $S$,
we can set
$$T_1 = (U\given S=s),\qquad T_2 = (V\given S=s), \qquad\hbox{and}\qquad T_3 = (W\given S=s)$$
and then average the above inequality over all $s\in \FF_2^n$, weighted by $\pr\{S=s\}$ to obtain
$$k\le \biggl(1+{\eta\over 3}\biggr)\bar\delta
+ {\eta\over 3}\sum_{i=1}^2\sum_{A\in\{U,V,W\}} \bigl(\dd\{X_i^*, A\given S\}-\dd\{X_i^*, X_i\}\bigr).$$
We now substitute the upper bounds~\refeq{\eqbardeltabound} and~\refeq{\eqdistancebounds} to get
$$\eqalign{
k &\le \biggl(1+{\eta\over 3}\biggr)\biggl(6\eta k - {1-5\eta \over 1-\eta}(2\eta k - I_1)\biggr)
+ {\eta\over 3}\biggl( (6-3\eta)k + 3(2\eta k - I_1)\biggr) \cr
&= (8\eta - \eta^2)k + \biggl(\eta-\Bigl(1+{\eta\over 3}\Bigr){1-5\eta\over 1-\eta} \biggr)
(2\eta k - I_1).\cr
}$$
With the choice $\eta = 1/9$,
$$\eta-\Bigl(1+{\eta\over 3}\Bigr){1-5\eta\over 1-\eta} = -{11\over 27} \le 0$$
and $2\eta k - I_1\ge 0$ by Lemma~{\lemionebound}, hence one concludes that
$$ k \le (8\eta + \eta^2)k = {73\over 81}k,$$
which is nonsense unless $k=0$.\slug
